SoK: The Evolution of Defenses Against Web Application Attacks

### 2014-11-11 docs/questioning.tex
more on discussion - ease of use

Such attacks include Cross-Site Scripting ({\sc xss})
attacks~\cite{SG07}, {\sc sql} injection attacks~\cite{RL12b},
Cross-Site Request Forgery ({\sc csrf}) attacks~\cite{LZRL09}, and
others. Attacks like these have been topping the vulnerability lists
of numerous bulletin providers like the Open Web Application Security
Project\footnote{\url{https://www.owasp.org/index.php/Category:OWASP_Top_Ten_Project}}
({\sc owasp}) and {\sc
  mitre}\footnote{\url{http://cwe.mitre.org/top25/}} for several
years. Consider the case of {\sc owasp}'s Top Ten project whose main
goal is to raise awareness about web application security by
identifying some of the most critical risks facing organizations. The
Top Ten project is popular, referenced by numerous researchers. In its
three consecutive Top Ten lists (2007, 2010, 2013), different
source code-driven injection attacks dominate the top five
positions. Malicious users find new ways to bypass defense mechanisms
by using a variety of techniques despite the numerous countermeasures
that are being introduced. Note that the number of systems developed
to counter {\sc sql} injection attacks until 2006 was more than
twenty~\cite{HVO06}. Since then the number has doubled. Researchers
have also indicated that the number of such attacks is steadily
increasing in recent years~\cite{SSL12}.

@@

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on source
code injection vulnerabilities. For every arrow there is a
corresponding tuple that includes a transition number for
every attack. Note that {\sc xss} attacks can also involve
phishing. Every {\it point} indicates a transition where mechanisms
detect attacks. For every point we provide the corresponding
references:
Point {\sc ub}~\cite{KJKV09,TNH07,NSS06,APKLM10,SPPGMMB04,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,GC09,VFJKKV07,SLMS14,BV08},
Point {\sc sb}~\cite{RDWDE07,JKK06a,JB07,NLC07,WPLKK09,JEP08},
Point {\sc eq}~\cite{BWS05,SW06,PS11,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
Point {\sc sdb}~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

As in medical diagnostic tests, application attack defense mechanisms
must demonstrate the presence of an attack. This, however, does not
make an application attack defense mechanism immediately useful. A
detection mechanism must be reliable. The accuracy of a detection
mechanism is gauged with the following
metrics~\cite{TDR2013,GFDLS06,A00}:
\begin{itemize}
\item {\bf Sensitivity}, the probability that an attack will be
  caught.
\item {\bf Specificity}, the probability that a normal interaction
  will test negative.
\item {\bf Positive Predictive Value} ({\sc ppv}), the probability that a
  reported attack is a real attack. It is the conditional probability
  that an event is an attack if the detection mechanism flags it as
  such. 
\item {\bf Negative Predictive Value} ({\sc npv}), the probability that if
  nothing is reported no attack has taken place. It is the conditional
  probability that an event is not an attack given that the detection
  mechanism flags it as normal.
\end{itemize}

@@

Sensitivity and specificity can be calculated based on test
data alone. To calculate the sensitivity, we run the test on a
controlled environment where we allow only attack events to reach the
system. The ratio of reported attacks over all attacks will give us
the sensitivity. Similarly, to calculate the specificity we can run
the test on a controlled environment where we allow only innocuous
events to reach the system. The ratio of non-reported events over all
events will give us the specificity.

@@

@@

Table~\ref{tab:comp} indicates that the authors of 3 out of 34
(8.8\%) publications, provided complete quadruples.
On the other hand, we see that 4 out of
34 (11.7\%) were not tested at all.
In this cases, all the four elements of the
quadruples are not available ({\sc na}).
% Maybe we should comment on the fact that these are
% frameworks coming from the etiological category
% and detect the attacks ``by design".
An interesting observation is that 10 out of 34
(29.41\%) of the authors performed only attacks
during their tests and did not test their mechanisms
to see if they produce any false alarms. The corresponding
quadruples in this case, are the ones that contain {\sc tp} and
{\sc fn} results but {\sc tn} and {\sc fp} results are not
available ({\sc na}). This stresses the fact that authors
maybe more interested into making sure that
their mechanisms can detect known attacks than seeing how
they respond under normal conditions.
In some occasions (where the number of {\sc tp} results
are not quantified), authors did not mention how many or
which attacks they performed.
Also, in some cases, we observe that even if
the authors report the existence of possible
false positive and negative results, they have not
quantified them ({\sc nq}).

@@

Consider the majority of mechanisms coming from
the policy enforcement category. In most cases
developers should modify multiple components
to use each mechanism. Specifically, mechanisms like
BrowserShield, and {\sc beep} require modifications
both on the server and the client side.
Thus, it would be difficult for them to be adopted by
both browser vendors and application developers.

@@

@@

Notes on taint tools:
\begin{enumerate}
\item The accuracy testing of {\sc csse}~\cite{PB05} was based on 
defects found in the security mailing
list: {\it Bugtraq}\footnote{\url{http://seclists.org/bugtraq/}}.
\item During their evaluation, the authors of~\cite{HCF05} have only performed
two attacks in one synthetic benchmark ({\sc owasp}'s
{\it WebGoat}\footnote{\url{https://www.owasp.org/index.php/Category:OWASP_WebGoat_Project}}).
\item The testing in~\cite{XBS06} was made based on
defects with specific {\sc cve id}s.
\item The authors of {\it {\sc wasc}}~\cite{NLC07}
mention the vulnerabilities (specific {\sc cve id}s)
exploited and the corresponding applications but they do not explicitly
state how many attacks they launched (``{\it [...] launched a variety of
attacks against it}").
\item The authors of~\cite{VFJKKV07} do the same.
However, even if they do not mention specific {\sc cve id}s,
they browsed applications without performing attacks.
\item Even if the authors of {\it Noxes}~\cite{KKVJ06,KJKV09} state in
both papers that ``{\it [...] we are planning to make the tool available as
a freeware utility}." the tool cannot be found anywhere on the Internet.
The vulnerabilities exploited during testing were reported
by {\it Bugtraq}\footnote{\url{http://seclists.org/bugtraq/}}.
\item {\it {\sc php} Aspis}~\cite{PMP11} was tested on only one
application (Wordpress 2.9.2) with multiple vulnerabilities (specific {\sc cve id}s).
\item During their initial tests, the authors of
managed to bypass the browser-based {\sc xss} filters of the
73\% of 1,602 real-world {\sc dom}-based {\sc xss} vulnerabilities.
Then they proposed an approach that detected all the attacks they
found (1169). This is a complex paper --- we need to elaborate.
\end{enumerate}

Notes on {\sc isr} mechanisms:
\begin{enumerate}
\item {\it {\sc sql}rand}~\cite{BK04}: Both exploits from {\sc cve} --- though
this is not mentioned in the publication.
\item {\it Noncespaces}~\cite{GC09} was only tested on one
known vulnerable application (TikiWiki). The authors mention that they performed
a number of attacks but they do not state how many.
\item The authors of {\it x{\sc js}}~\cite{APKLM10}, have tested
their mechanisms based on real-world attacks coming from
{\sc xss}ed.com.\footnote{www.xssed.com}
\end{enumerate}

Notes on policy enforcement tools:
\begin{enumerate}
\item Just like the authors of {\it x{\sc js}~\cite{APKLM10},
the authors of {\sc dsi}}~\cite{NSS06}, have tested
their mechanisms based on real-world attacks coming from
{\sc xss}ed.com.\footnote{www.xssed.com}
\item The testing of {\it BrowserShield}~\cite{RDWDE07}
involved {\sc xss} vulnerabilities reported by Microsoft
in 2005.\footnote{\url{https://technet.microsoft.com/en-us/security/bulletin}}
The link currently shows defects for 2014.
\item {\it CoreScript}~\cite{YCIS07} is more of a programming
language\footnote{``[...] uses program instrumentation for
JavaScript where untrusted JavaScript
code goes through a rewriting process according to a security policy
before executing them in the browser"~\cite{PSC09}.}
than an {\sc IDS}.
\item Surprisingly, {\sc met}~\cite{ELX07} was not tested at all.
\item {\sc beep}~\cite{TNH07}'s test suite was based on 61 {\sc xss}
attack vectors published by \url{ha.ckers.org}.
\item {\sc soma}~\cite{OWVS08} --- all synthetics.
\item The attacks used for the testing of {\it Blueprint}~\cite{LV09},
came from \url{ha.ckers.org}.
\item {\it Phung et al.}~\cite{PSC09} also utilized \url{ha.ckers.org}.
\item {\it WebJail}~\cite{VDDPJ11} --- not tested at all.
\item {\it ConScript}~\cite{ML10} --- not tested at all. A scheme
similar to {\it CoreScript}~\cite{YCIS07}.
\item {\it j{\sc csrf}}~\cite{PS11} --- based on {\sc csrf}
defects with specific {\sc cve id}s.
\item The authors of {\it TreeHouse}~\cite{IW12} have tested
only the overhead and how easy is to incorporate the mechanism
into an application.
\end{enumerate}

Notes on training-based mechanisms:
\begin{enumerate}
\item {\it {\sc didafit}}~\cite{LLW02} is one of
the first {\sc ids}.
\item {\sc amnesia}~\cite{HO05,HO06,HO05b} was tested
on applications with known vulnerabilities taken
from \url{http://gotocode.com} plus two synthetic benchmarks.
Interestingly, in one of the papers~\cite{HO06},
the authors mention that the tool can produce both {\sc fp} and
{\sc fn}.
\item {\it libAnomaly}~\cite{VMV05} --- one application tested ({\sc php}-Nuke).
\item {\sc sqlg}uard~\cite{BWS05} --- no testing at all.
\item {\sc sm}ask~\cite{JB07} was tested on real-world applications
with exploits from {\sc cve} --- though this is not mentioned in the publication.
\item {\sc xssds}~\cite{JEP08} --- testing was based on 
defects found in the security mailing
list: {\it Bugtraq}\footnote{\url{http://seclists.org/bugtraq/}}.
The authors provide multiple {\sc fp} rates depending on the experiment.
\item {\sc xss-guard}~\cite{BV08} was tested on real-world applications
with exploits from {\sc cve}.
\item {\sc swap}~\cite{WPLKK09} was tested on a real-world application ({\sc phpbb})
with exploits from {\sc cve} --- though this is not mentioned in the publication.
%\item {\it Laranjeiro et al.}~\cite{LVM09,ALVM09,LVM10} have developed a
%mechanism to detect {\sc sql} and {\sc xp}ath injection attacks
%only on web services. Their testing involve a real-world benchmark
%({\sc tpc}-app).\footnote{\url{http://www.tpc.org/}}
\item {\it Diglossia}~\cite{SMS13} was tested on real-world applications
with exploits from {\sc cve}.
\end{enumerate}

Notes:
\begin{enumerate}
\item Note that most mechanisms that implement etiological
approaches have not been extensively tested in terms of
accuracy.
\item We need to mention that there are mechanisms that
do not need to be established purely through testing
Since they provide systematic arguments as to why
the design is secure against attacks.
\item Policy enforcement is the
only approach that is
used to prevent {\sc csrf} attacks.
\item See j{\sc csrf} for {\sc soma}'s drawbacks.
\item Discuss about how tools can be easily modified
to detect other attacks ({\sc sql}rand).
\end{enumerate}

@@

>>>>

Such attacks include Cross-Site Scripting ({\sc xss})
attacks~\cite{SG07}, {\sc sql} injection attacks~\cite{RL12b},
Cross-Site Request Forgery ({\sc csrf}) attacks~\cite{LZRL09}, and
others. Attacks like these have been topping the vulnerability lists
of numerous bulletin providers like the Open Web Application Security
Project\footnote{\url{https://www.owasp.org/index.php/Category:OWASP_Top_Ten_Project}}
({\sc owasp}) and {\sc
  mitre}\footnote{\url{http://cwe.mitre.org/top25/}} for several
years. Consider the case of {\sc owasp}'s Top Ten project whose main
goal is to raise awareness about web application security by
identifying some of the most critical risks facing organizations. The
Top Ten project is popular, referenced by numerous researchers. In its
three consecutive Top Ten lists (2007, 2010, 2013), different
source code-driven injection attacks dominate the top five
positions. Malicious users find new ways~\cite{HNSHS12,DKH14}
to bypass defense mechanisms
by using a variety of techniques despite the numerous countermeasures
that are being introduced. Note that the number of systems developed
to counter {\sc sql} injection attacks until 2006 was more than
twenty~\cite{HVO06}. Since then the number has doubled. Researchers
have also indicated that the number of such attacks is steadily
increasing in recent years~\cite{SSL12}.

@@

\begin{figure*}
\begin{center}
\leavevmode
\includegraphics[scale=0.61]{attacks-steps.pdf}
\end{center}
\caption{\label{fig:attacks}Attack model demonstrating web
application attacks based on source
code injection vulnerabilities. For every arrow there is a
corresponding tuple that includes a transition number for
every attack. Every {\it point} indicates a transition where mechanisms
detect attacks. For every point we provide the corresponding
references:
Point {\sc ub}~\cite{KJKV09,TNH07,NSS06,APKLM10,SPPGMMB04,ML10,YCIS07,PSC09,VDDPJ11,OWVS08,DDHPJ10,GC09,VFJKKV07,SLMS14,BV08},
Point {\sc sb}~\cite{RDWDE07,JKK06a,JB07,NLC07,WPLKK09,JEP08},
Point {\sc eq}~\cite{BWS05,SW06,PS11,HCF05,XBS06,PB05,PMP11,MS09,HO05,SMS13},
Point {\sc sdb}~\cite{BK04,LLW02,VMV05}.}
\end{figure*}

@@

Application attack defense mechanisms
must demonstrate the presence of an attack. This, however, does not
make an application attack defense mechanism immediately useful. A
detection mechanism must be reliable. The accuracy of a detection
mechanism is gauged with the following
metrics~\cite{TDR2013,GFDLS06,A00}:
\begin{itemize}
\item {\bf Sensitivity}, the probability that an attack will be
  caught.
\item {\bf Specificity}, the probability that a normal interaction
  will test negative.
\item {\bf Positive Predictive Value} ({\sc ppv}), the probability that a
  reported attack is a real attack. It is the conditional probability
  that an event is an attack if the detection mechanism flags it as
  such. 
\item {\bf Negative Predictive Value} ({\sc npv}), the probability that if
  nothing is reported no attack has taken place. It is the conditional
  probability that an event is not an attack given that the detection
  mechanism flags it as normal.
\end{itemize}

@@

\noindent
Sensitivity and specificity can be calculated based on test
data alone. To calculate the sensitivity, we run the test on a
controlled environment where we allow only attack events to reach the
system. The ratio of reported attacks over all attacks will give us
the sensitivity. Similarly, to calculate the specificity we can run
the test on a controlled environment where we allow only innocuous
events to reach the system. The ratio of non-reported events over all
events will give us the specificity.

@@

\subsection{Ease of Use}
\label{sec:deployment}

@@

Table~\ref{tab:comp} indicates that the authors of 3 out of 34
(8.8\%) publications, provided complete quadruples.
On the other hand, we see that 4 out of
34 (11.7\%) were not tested at all.
In this cases, all the four elements of the
quadruples are not available ({\sc na}).
An interesting observation is that 10 out of 34
(29.41\%) of the authors performed only attacks
during their tests and did not test their mechanisms
to see if they produce any false alarms. The corresponding
quadruples in this case, are the ones that contain {\sc tp} and
{\sc fn} results but {\sc tn} and {\sc fp} results are not
available ({\sc na}). This stresses the fact that authors
maybe more interested into making sure that
their mechanisms can detect known attacks than seeing how
they respond under normal conditions.
In some occasions (where the number of {\sc tp} results
are not quantified), authors did not mention how many or
which attacks they performed.
Also, in some cases, we observe that even if
the authors report the existence of possible
false positive and negative results, they have not
quantified them ({\sc nq}).

@@

As we mentioned in Subsection~\ref{sec:deployment},
there are various reasons why a mechanism could
be difficult to deploy.

Consider the majority of mechanisms coming from
the policy enforcement subcategory. In most cases
developers should modify multiple components
to use each mechanism. Specifically, mechanisms like
BrowserShield, and {\sc beep} require modifications
both on the server and the client side.
Thus, it would be difficult for them to be adopted by
both browser vendors and application developers.
On the other hand, there are cases where the policies
introduced in the server, are enforced on the client-side,
via a library embedded in the server's response
(i.e. Blueprint). Such an approach is really convenient
since no modifications are needed on the client.

Extensive modifications in the application's source
code can also be a reason that can make a mechanism
difficult to use. {\sc sql}rand, {\sc amnesia}
and mechanisms coming from the taint tracking
subcategory are such examples.
In the first two cases, programmers should modify every
code fragment that involves the execution of a query.
In the latter case they should also change
all the code fragments that involve user-input handling. 
In addition, there are cases (i.e. ConScript and the system
by Phung et al.) that require many modifications
by the developers in order to introduce their policies.
Note also that in many cases, developers are required to
learn a new {\sc api}.

Finally, a disadvantage of mechanisms that involve training
({\sc sd}river, {\sc xss-guard}), is that when code statements
are altered, a new training phase is necessary. However,
with the increased adoption of the use of automated testing
and continuous integration frameworks, this phase could be
easily repeated.

@@

Another interesting finding involves the computational overhead.
We observe that the average overhead that can be extracted from
the publications that provide the overhead in percentage terms is
18\%. Relative research on protection mechanisms~\cite{SPWS13}
has indicated that mechanisms that introduce an overhead
larger than 10\% do not tend to gain wide adoption in
production environments which in this case could be a
reason why some these tools have not been adopted.

Cite~\cite{DKH14}.

@@

@@

We believe that forming a
standard way to test such mechanisms in terms
of accuracy and computational performance
would be one of the reasons of adopting them.
